Bounds on Coding Theory from Algebraic Geometry

### 2021-04-05 draft.tex
Updating preamble to match Haverford's guideline

\documentclass[11pt, letterpaper]{amsart}
%\usepackage{amsaddr}

\usepackage[isbn=false,url=false,doi=false,style=alphabetic]{biblatex} 
\addbibresource{sample.bib}

\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}
\setlist{itemsep=.5em, parsep = .5em} % < Usually a good idea

\newtheorem{thm}[]{Theorem}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[]{Definition}

@@

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\Ima}{Im}

\title{Bounds on Coding Theory from Algebraic Geometry}
\author{Guilherme Zeus Dantas e Moura}
%\address{Haverford College}
%\email{gdantasemo@haverford.edu}
%\date{Spring, 2021. Last change in \today}
\date{last compiled \today}

@@

\maketitle
\begin{abstract}    
    Coding theory is concerned with finding efficient ways to encode a message so that one may correct errors in the message. In algebraic coding theory, we study efficient codes generated from algebraic geometric methods.

@@

\begin{defn}[Hamming distance]
    We define \emph{Hamming distance} between $\mathbf{x} = (x_1, \dots, x_n), \mathbf{y} = (y_1, \dots, y_n) \in A^n$ as
    \[ \dist(\mathbf{x}, \mathbf{y}) = \#\left( x_i \neq y_i \mid i \in \{1, 2, \dots, n\} \right),\]
    in other words, the number of positions $\mathbf{x}$ and $\mathbf{y}$ differ.
\end{defn}

@@

\begin{defn}[Parameters of a code]
    If $C$ is a linear code over $A$, we define \emph{dimension of $C$} as $k = \dim_A(C)$ and \emph{minimum distance of $C$} as $d = \min\left\{\dist(\mathbf{x}, \mathbf{y}) \mid \mathbf{x}, \mathbf{y} \in C\right\}$. (\emph{If $C$ is a nonlinear code with length $n$, we can coherently define $k = \log_n{\left|C\right|}$.}) The length $n$, dimension $k$ and minimum distance $d$ are the \emph{parameters} of $C$.
\end{defn}

@@

\section{Singleton bound and a promising example} \label{s:singleton}
\begin{thm}[Singleton Bound]\label{thm:singleton_bound}
If $C$ is a code with parameters $n, k, d$, then \[k + d \le n + 1,\] or equivalently, \[R + \delta \le 1 + 1/n.\]
\end{thm}

\begin{proof}
	We will provide the proof for Theorem \ref{thm:singleton_bound} when $C$ is a linear code.
\end{proof}

\begin{defn}[Reed--Solomon Codes]
	Let $q$ be a power of a prime, and $\mathbb{F}_q = \{\alpha_1, \alpha_2, \dots, \alpha_q\}$ the field with $q$ elements. Let $k$ be an integer, and $\mathcal{L}_k$ the set of all polylomials over $\mathbb{F}_q$ with degree smaller than $k$. Let $k \le n \le q$ be an integer. The Reed--Solomon code $RS_q(n, k)$ over $\mathbb{F}_q$ is \[
		RS_q(n, k) = \left\{\left(f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)\right) \mid f \in \mathcal{L}_k\right\}.
	\]
\end{defn}

@@

\begin{proof}
	$R_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it has length $n$.  
	Note that $\mathcal L_k$ is a vector space over $\mathbb{F}_q$. Note that $\{1, x, x^2, \dots, x^{k-1}\}$ is a choice of basis for this vector space, thus it has dimension $k$. Consider the map $\phi: \mathcal L_k \to \mathbb{F}_q^n$ given by \[
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)).
	\]

	Note that the map $\phi$ is a linear transformation. Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polylomial, which implies $f = g$. Therefore, $\phi$ is also injective. This implies that the dimension of the domain $\mathcal L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dim RS_q(n, k) = k$.

	Finally, consider distinct $f, g \in \mathcal L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$. Futhermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots. Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	Thus, when $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which along the \nameref{thm:singleton_bound} implies \[
		k + d_\text{min} = n + 1.
	\]

@@

\section{Generalized Reed--Solomon codes} \label{s:grs}
We shall redefine the Reed--Solomon codes using language related to a projective line. There is a way to replace the ``projective line'' with a ``projective plane curve'' and create other codes, called \emph{Generalized Reed--Solomon codes} or simply \emph{algebraic geometric codes}. We want large $R$ and $\delta$, and these codes yield \begin{equation}\label{eq:g/n} R + \delta \ge 1 + 1/n - g/n,\end{equation} where $n$ is the number of rational points of a curve $X$, with genus $g$.

>>>>

\documentclass[11pt]{amsart}
\usepackage[letterpaper,left=125pt,right=125pt]{geometry}

\usepackage[english]{babel}
\usepackage{amsmath, amssymb, amsfonts, amsthm, graphicx}

\usepackage[colorlinks,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage[backend=biber,isbn=false,url=false]{biblatex} 
\addbibresource{sample.bib}

\theoremstyle{theorem}
\newtheorem{thm}{Theorem}[section]

\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{alg}[thm]{Algorithm}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{conj}[thm]{Conjecture}

\theoremstyle{example}
\newtheorem{prob}[thm]{Problem}

\theoremstyle{remark}
\newtheorem{exmp}[thm]{Example}
\newtheorem{rem}[thm]{Remark}
\newtheorem{claim}[thm]{Claim}  
\renewcommand{\theclaim}{}

\numberwithin{equation}{section}

\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{enumitem}
\setlist{itemsep=.5em, parsep = .5em}
\usepackage{lipsum}

@@

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\Ima}{Im}
\DeclareMathOperator{\dimension}{dim}

\title[Codes from Algebraic Geometry]{Bounds on Coding Theory from Algebraic Geometry}
\author{Guilherme Zeus Dantas e Moura}
\date{last compiled \today} % <---- Change this date to the format YYYY-MM-DD

@@

\maketitle
\begin{abstract} % NOT REQUIRED
    Coding theory is concerned with finding efficient ways to encode a message so that one may correct errors in the message. In algebraic coding theory, we study efficient codes generated from algebraic geometric methods.

@@

\begin{defn}[Hamming distance]
    We define \emph{Hamming distance} between $\mathbf{x} = (x_1, \dots, x_n), \mathbf{y} = (y_1, \dots, y_n) \in A^n$ as
    \begin{equation}
		\dist(\mathbf{x}, \mathbf{y}) = \#\left( x_i \neq y_i \mid i \in \{1, 2, \dots, n\} \right),
	\end{equation}
    in other words, the number of positions $\mathbf{x}$ and $\mathbf{y}$ differ.
\end{defn}

@@

\begin{defn}[Parameters of a code]
    If $C$ is a linear code over $A$, we define \emph{dimension of $C$} as $k = \dimension_A(C)$ and \emph{minimum distance of $C$} as $d = \min\left\{\dist(\mathbf{x}, \mathbf{y}) \mid \mathbf{x}, \mathbf{y} \in C\right\}$. (\emph{If $C$ is a nonlinear code over an alphabet with size $q$, we can coherently define $k = \log_q{\left|C\right|}$.}) The length $n$, dimension $k$ and minimum distance $d$ are the \emph{parameters} of $C$.
\end{defn}

@@

\section{Singleton bound and a promising example} \label{s:singleton}
\begin{thm}[Singleton Bound]\label{thm:singleton_bound}
If $C$ is a code with parameters $n, k, d$, then 
	\begin{equation}
		k + d \le n + 1,
	\end{equation} 
	or equivalently, 
	\begin{equation}
		R + \delta \le 1 + 1/n.
	\end{equation}
\end{thm}

\begin{proof}
	We will provide the proof for Theorem \ref{thm:singleton_bound} when $C$ is a linear code.

	WRITE PROOF.
\end{proof}

\begin{defn}[Reed--Solomon Codes]
	Let $q$ be a power of a prime, and $\mathbb{F}_q = \{\alpha_1, \alpha_2, \dots, \alpha_q\}$ the field with $q$ elements. Let $k$ be an integer, and $\mathcal{L}_k$ the set of all polylomials over $\mathbb{F}_q$ with degree smaller than $k$. Let $k \le n \le q$ be an integer. The Reed--Solomon code $RS_q(n, k)$ over $\mathbb{F}_q$ is 
	\begin{equation}
		RS_q(n, k) = \left\{\left(f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)\right) \mid f \in \mathcal{L}_k\right\}.
	\end{equation}
\end{defn}

@@

\begin{proof}
	$R_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it has length $n$.  
	Note that $\mathcal L_k$ is a vector space over $\mathbb{F}_q$. Note that $\{1, x, x^2, \dots, x^{k-1}\}$ is a choice of basis for this vector space, thus it has dimension $k$. Consider the map $\phi: \mathcal L_k \to \mathbb{F}_q^n$ given by 
	\begin{equation}
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)).
	\end{equation}

	Note that the map $\phi$ is a linear transformation. Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polylomial, which implies $f = g$. Therefore, $\phi$ is also injective. This implies that the dimension of the domain $\mathcal L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dimension RS_q(n, k) = k$.

	Finally, consider distinct $f, g \in \mathcal L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$ roots. Futhermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots. Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	If we choose $f, g$ such that $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which together with \nameref{thm:singleton_bound} implies 
	\begin{equation}
		k + d_\text{min} = n + 1.
	\end{equation}

@@

\section{Generalized Reed--Solomon codes} \label{s:grs}
We shall redefine the Reed--Solomon codes using language related to a projective line. There is a way to replace the ``projective line'' with a ``projective plane curve'' and create other codes, called \emph{Generalized Reed--Solomon codes} or simply \emph{algebraic geometric codes}. We want large $R$ and $\delta$, and these codes yield 
	\begin{equation}
		\label{eq:g/n} R + \delta \ge 1 + 1/n - g/n,
	\end{equation}
	where $n$ is the number of rational points of a curve $X$, with genus $g$.
